---
title: "677 midterm"
output: pdf_document
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amsfonts}
---
# order statistic

## Uniform Distribution

Consider the uniform distribution $U(a,b)$. The PDF and CDF are given by:

- PDF: $f(x) = \frac{1}{b-a}$, for $a \leq x \leq b$
- CDF: $F(x) = \frac{x-a}{b-a}$, for $a \leq x \leq b$

For the $k$-th order statistic $X_{(k)}$ from a sample of size $n$:

- PDF: $f_{(k)}(x) = \frac{n!}{(k-1)!(n-k)!} [F(x)]^{k-1}[1 - F(x)]^{n-k} f(x)$

Applying the CDF and PDF of the uniform distribution:

- PDF of $X_{(k)}$: $f_{(k)}(x) = \frac{n!}{(k-1)!(n-k)!} \left(\frac{x-a}{b-a}\right)^{k-1} \left(1 - \frac{x-a}{b-a}\right)^{n-k} \frac{1}{b-a}$

## Exponential Distribution

For the exponential distribution with rate $\lambda$, the PDF and CDF are:

- PDF: $f(x) = \lambda e^{-\lambda x}$, for $x \geq 0$
- CDF: $F(x) = 1 - e^{-\lambda x}$, for $x \geq 0$

The PDF of the $k$-th order statistic $X_{(k)}$:

- PDF of $X_{(k)}$: $f_{(k)}(x) = \frac{n!}{(k-1)!(n-k)!} [1 - e^{-\lambda x}]^{k-1} e^{-\lambda x[n-k]} \lambda e^{-\lambda x}$

## Normal distribution

For the normal distribution, deriving the distribution of order statistics analytically is more complex due
to the non-linearity of its CDF. Instead, we often use numerical methods or simulations to study the order
statistics of normal samples

## Simulation of uniform

In this chapter,I choose s order statistics that is:0th,25th,50th,75th,100th 
```{r,echo=FALSE}
# Set seed for reproducibility
set.seed(123)

# Sample size and number of simulations
n <- 1000
num_simulations <- 10000

# Generate samples from uniform distribution
uniform_samples <- matrix(runif(n * num_simulations), ncol = n)

# Calculate order statistics for each simulation
uniform_order_stats <- apply(uniform_samples, 1, function(x) quantile(x, probs = c(0, 0.25, 0.5, 0.75, 1)))

# Plot histograms for the order statistics
par(mfrow = c(3, 2))  # Arrange plots in 3 rows, 2 columns
for (i in 1:5) {
  hist(uniform_order_stats[i, ], main = paste("Uniform Distribution - Order Statistic", i), xlab = "Value", breaks = 40, probability = TRUE)
}


```

## Simulation of exponetianal

```{r,echo=FALSE}
# Generate samples from exponential distribution
lambda <- 1
exponential_samples <- matrix(rexp(n * num_simulations, rate = lambda), ncol = n)

# Calculate order statistics for each simulation
exponential_order_stats <- apply(exponential_samples, 1, function(x) quantile(x, probs = c(0, 0.25, 0.5, 0.75, 1)))

# Plot histograms for the order statistics
for (i in 1:5) {
  hist(exponential_order_stats[i, ], main = paste("Exponential Distribution - Order Statistic", i), xlab = "Value", breaks = 40, probability = TRUE)
}

```
## Simulation of normal

```{r,echo=FALSE}
# Generate samples from normal distribution
normal_samples <- matrix(rnorm(n * num_simulations), ncol = n)

# Calculate order statistics for each simulation
normal_order_stats <- apply(normal_samples, 1, function(x) quantile(x, probs = c(0, 0.25, 0.5, 0.75, 1)))

# Plot histograms for the order statistics
for (i in 1:5) {
  hist(normal_order_stats[i, ], main = paste("Normal Distribution - Order Statistic", i), xlab = "Value", breaks = 40, probability = TRUE)
}

```

